---
title: "Mediaflux MPP Metadata"
author: "Jessica Chung"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
    toc_depth: 4
    code_folding: show
---

From tidy data generated from `metadata_wrangling.Rmd`, generate csv files
to feed into Mediaflux via a Python script.

Create an asset for each possum. Associate the asset to the sequencing files 
with a relationship. Each possum asset should have a variable number of metadata 
libraries (correlated to the number of times it's been captured).

-----

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=TRUE, message=TRUE, error=TRUE, echo=TRUE)
knitr::opts_knit$set(root.dir = "..")
options(digits=6)
options(width=160)
```

```{r message=FALSE}
library("ProjectTemplate")
load.project()
```

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(stringr)
library(readxl)
library(forcats)
```

```{r}
append_note <- function(x, existing_value, sep="; ") {
  # Append a note separated with a delimiter if a comment already exists.
  # Else, replace NA with note.
  joined_values <- paste(existing_value, x, sep=sep)
  new_value <- ifelse(is.na(existing_value) | existing_value == "",
                      x, joined_values) %>%
    str_replace("; $", "")
  return(new_value)
}
```

-----

# Load data

```{r}
tidy_data <- readRDS(file="cache/tidy_data.rds")
microsat_data <- readRDS(file="cache/microsat_data.rds")
replaced_tags <- readRDS(file="cache/replaced_tags.rds")
```

```{r include=FALSE, eval=FALSE}
tidy_data %>% View()
```


-----

# Static information

Each possum (or bandicoot, in the future) asset should have one of these rows
which translates into a metadata library.

```{r}
# Information relevant to sample (should be the same for each capture instance)
# Individuals sometimes have different mountains and wild_captive status at different times
x <- tidy_data %>% select(id, sex, country:state_or_region, taxon_id:common_name)

# Check that rows are the same for each ID
stopifnot(x %>% unique %>% group_by(id) %>% summarise(n=n()) %>% 
            filter(n != 1) %>% nrow %>% magrittr::equals(0))

static_df <- x %>% unique
static_cols <- colnames(x)
```

```{r}
static_df %>% head
dim(static_df)

stopifnot(length(static_df$id) == length(unique(static_df$id)))
```

-----

# Variable informaiton

Values can change with each capture record. Each possum asset can have zero or
multiple of these.

```{r}
# Information that can change with each capture
variable_cols <- c("id", "sex", colnames(tidy_data)[! colnames(tidy_data) %in% static_cols])
variable_df <- tidy_data[,variable_cols] %>% 
  select(id, date, location, sex, everything())
```

```{r}
variable_df %>% head
```


### Special records

These are either captive from zoos or candidate males from Hotham, Timms Spurs.
These records don't have dates (but have years), location (but has mountain instead),
life_stage.

```{r}
# variable_df %>% filter(is.na(date))

# Remove location field since the mountain field contains the same information.
special_records <- variable_df %>% filter(is.na(date)) %>%
  select(-date, -location, -life_stage, -identified_by, -collector) %>%
  select(id, year, mountain, everything())

# Note: even though timms spur males have testis size, not classified as enlarged 
# because they aren't explicitly labeled adult

# Rename mountain to source and add 'Healesville Sanctuary' as a value.
special_records <- special_records %>% rename(source=mountain) %>%
  mutate(source=ifelse(! is.na(source_population) & 
                         str_detect(source_population, 'Healesville'), 
                       "Healesville Sanctuary zoo release", source)) %>%
  select(-source_population)
```

```{r}
special_records %>% head
```


### Wild capture records

```{r}
# Wild capture records always have dates recorded
wild_capture_records <- variable_df %>% filter(! is.na(date))

# Check columns don't hold useful information
stopifnot(wild_capture_records$wild_captive == "wild")
stopifnot(is.na(wild_capture_records$source_population))

# Remove unneeded columns
wild_capture_records <- wild_capture_records %>% 
  select(-mountain, -year, -wild_captive, -source_population)

```

```{r}
# What's the max number of captures for any particular possum?
wild_capture_records %>% group_by(id) %>%
  summarise(n=n()) %>% arrange(desc(n)) %>% head
```

```{r}
# wild_capture_records %>% filter(id == "0241")
```

-----

# Genotype records

One possum asset can have multiple microsat records. Ideally, a possum should
only have one, but some possums have been genotyped multiple times and have
non-concordant genotypes.

```{r}
microsat_data %>% head
```

-----

# Summary metadata

If useful, can also store a summary of the wild capture metadata.
E.g. earliest capture, juvenile capture, adult capture, etc.

```{r}
last_non_NA <- function(x) {
  last(x[!is.na(x)])
}
```

```{r}
summary_data <- wild_capture_records %>%
  group_by(id) %>%
  summarise(earliest_capture=min(date),
            latest_capture=max(date),
            number_of_captures=n(),
            juvenile_capture="juvenile" %in% life_stage,
            immature_capture="immature adult" %in% life_stage,
            adult_capture="adult" %in% life_stage,
            most_recent_weight=last_non_NA(weight),
            most_recent_head=last_non_NA(head),
            most_recent_body=last_non_NA(body),
            most_recent_tail=last_non_NA(tail))
```

```{r}
summary_data %>% head
```

```{r}
summary_data %>% filter(juvenile_capture, adult_capture) %>%
  arrange(desc(number_of_captures)) %>% head
```

```{r include=FALSE, eval=FALSE}
# wild_capture_records %>% filter(id == "0577") %>%
#   arrange(date) %>% View()
```

-----

# Replaced tags

```{r}
static_df$alternative_id = NA
for (tag in names(replaced_tags)) {
  print(tag)
  static_df <- static_df %>% 
  mutate(alternative_id=ifelse(id == tag, 
                               paste(replaced_tags[[tag]], collapse=","),
                               alternative_id))
}
```

-----

# Save files

```{r}
write.table(static_df, file="output/general.tsv", sep="\t", 
            row.names=FALSE, quote=FALSE)
write.table(special_records, file="output/special_records.tsv", sep="\t", 
            row.names=FALSE, quote=FALSE)
write.table(wild_capture_records, file="output/wild_capture_records.tsv", sep="\t", 
            row.names=FALSE, quote=FALSE)
write.table(microsat_data, file="output/microsat_data.tsv", sep="\t", 
            row.names=FALSE, quote=FALSE)
write.table(summary_data, file="output/summary_data.tsv", sep="\t", 
            row.names=FALSE, quote=FALSE)
```

-----

# Session Info

```{r}
devtools::session_info()
```


