---
title: "MPP Metadata Wrangling"
author: "Jessica Chung"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
    toc_depth: 4
    code_folding: show
---

Organise Mountain Pygmy Possum sample metadata from Excel spreadsheets.

-----

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=TRUE, message=TRUE, error=TRUE, echo=TRUE)
knitr::opts_knit$set(root.dir = "..")
options(digits=6)
options(width=160)
```

```{r message=FALSE}
library("ProjectTemplate")
load.project()
```

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(stringr)
library(readxl)
library(forcats)
```

```{r}
append_note <- function(x, existing_value, sep="; ") {
  # Append a note separated with a delimiter if a comment already exists.
  # Else, replace NA with note.
  joined_values <- paste(existing_value, x, sep=sep)
  new_value <- ifelse(is.na(existing_value) | existing_value == "",
                      x, joined_values) %>%
    str_replace("; $", "")
  return(new_value)
}
```

```{r include=FALSE}
stopifnot(append_note(c("a","b"), c(NA, "1")) ==
            c("a", "1; b"))
```

------

# Read Spreadsheets

Capture file stores instances of possum capture.  
Genotype file stores microsat data and other possum data.  

```{r}
# Filenames
capture_filename <- "data/Btr2016_Master_25July2017_FINAL.xlsx"
genotype_filename <- "data/MPP_Master_Genotype_Phenotype_March2017 copy.xlsx"
```

Read in capture spreadsheet.

```{r}
capture_xlsx <- read_excel(path=capture_filename, col_names=FALSE) %>%
  as.data.frame()

# Rename columns and rows
new_col_names <- capture_xlsx[1,] %>% str_replace("1st wt", "first_weight") %>% 
  str_replace_all(" ", "_") %>% tolower()
data <- capture_xlsx[-1,] %>% magrittr::set_colnames(new_col_names) %>%
  magrittr::set_rownames(NULL) %>% filter(! is.na(date))
```

View capture spreadsheet data.

```{r}
dim(data)
data %>% head
```

Read in genotype spreadsheet.

```{r}
years <- c(2010, rep(2011:2015, each=2))
sheet_names <- c(paste(years, c("", " juv"), sep=""), 
                 "Zoo release 2013", "Hotham males 2010", 
                 "Hotham males 2011", "Timms Spur males 2014")
all_sheets <- list()

for (sheet in sheet_names) {
  # print(sheet)
  temp_sheet <- read_excel(path=genotype_filename, sheet=sheet, 
                           col_names=TRUE)
  sheet_split <- strsplit(sheet, " ")
  temp_sheet$from_sheet <- sheet
  # Juveniles records are stored in sheets ending with 'juv'
  if (str_detect(sheet, "juv$")) {
    temp_sheet$life_stage <- "juvenile"
  } else {
    temp_sheet$life_stage <- "post-juvenile adult"
  }
  
  # Insert specific data for each sheet
  if (sheet == "Zoo release 2013") {
    temp_sheet$YEAR <- "2013"
    temp_sheet$wild_captive <- "captive"
    temp_sheet$MOUNTAIN <- NA
    temp_sheet$SOURCE_POPULATION <- 
      "Australia:Victoria, Mount Buller, captive-bred at Healesville Sanctuary"
  }
  if (sheet == "Hotham males 2010") {
    temp_sheet$YEAR <- "2010"
    temp_sheet$SEX <- "2"
    temp_sheet$MOUNTAIN <- "Mount Hotham"
    temp_sheet$wild_captive <- "wild"
  }
  if (sheet == "Hotham males 2011") {
    temp_sheet$YEAR <- "2011"
    temp_sheet$MOUNTAIN <- "Mount Hotham"
    temp_sheet$wild_captive <- "wild"
  }
  if (sheet == "Timms Spur males 2014") {
    temp_sheet$YEAR <- "2014"
    temp_sheet$MOUNTAIN <- "Timms Spur"
    temp_sheet$wild_captive <- "wild"
  }
  if (sheet %in% sheet_names[1:11]) {
    temp_sheet$YEAR <- sheet_split[[1]][1]
    temp_sheet$MOUNTAIN <- "Mount Buller"
    temp_sheet$wild_captive <- "wild"
  }

  # Change duplicate column names to be unique
  is_dup <- temp_sheet %>% colnames %>% duplicated
  index <- seq_len(sum(is_dup))
  colnames(temp_sheet)[is_dup] <- paste(colnames(temp_sheet)[is_dup], index, sep="_")
  
  # Get columns which are factors
  is_factor <- lapply(temp_sheet, class) %>% unlist %>% `==`("factor")
  
  # Convert factor columns into characters
  if (sum(is_factor) > 0) {
    message(paste("Converting factors into characters: ", names(is_factor)[is_factor]))
    temp_sheet <- temp_sheet %>% mutate_if(is.factor, as.character)
  }
  
  # Identify columns containing microsat genotypes, edit names appropriately,
  # and add names for columns containing second allele
  microsat_col1s <- which(!is.na(str_match(colnames(temp_sheet), pattern = "B[PC]")))
  for(i in microsat_col1s){
    col1_name <- colnames(temp_sheet)[i]
    col2_name <- colnames(temp_sheet)[i+1]
    if(str_sub(col1_name, -1, -1)==")"){
      new_col1_name <- paste(str_sub(col1_name, 1, -1), "allele1", sep="") %>%
        str_replace_all("[()]", "_")
      new_col2_name <- paste(str_sub(col1_name, 1, -1), "allele2", sep="") %>%
        str_replace_all("[()]", "_")
      colnames(temp_sheet)[i:(i+1)] <- c(new_col1_name, new_col2_name)
    }
  }
  # temp_sheet <- select(temp_sheet, -starts_with("X"))

  # Store data
  all_sheets[[sheet]] <- temp_sheet
}
```

Sanity check before merging data frames.

```{r message=FALSE}
# Common columns. All dataframes should have these columns
common_columns <-
  c("SITE", "DATE", "DAY", "LINE", "POINT", "SPECIES", "ID", "SEX", 
    "AGE", "PY", "WEIGHT", "NEW/RE", "Repro Cond", "Tail", "Body", 
    "Head", "First captured", "Retag", "Notes", "hybrid alleles", 
    "BP1_FAM_allele1", "BP1_FAM_allele2", "BP2_FAM_allele1", "BP2_FAM_allele2", 
    "BP3_NED_allele1", "BP3_NED_allele2", "BP11_FAM_allele1", "BP11_FAM_allele2", 
    "BP12_VIC_allele1", "BP12_VIC_allele2", "BP13_NED_allele1", "BP13_NED_allele2", 
    "BP17_VIC_allele1", "BP17_VIC_allele2", "BP20_NED_allele1", "BP20_NED_allele2", 
    "BC1_FAM_allele1", "BC1_FAM_allele2", "BC4_FAM_allele1", "BC4_FAM_allele2", 
    "BC6_FAM_allele1", "BC6_FAM_allele2", "BC11_VIC_allele1", "BC11_VIC_allele2", 
    "BC13_VIC_allele1", "BC13_VIC_allele2", "BC14_VIC_allele1", "BC14_VIC_allele2", 
    "BC18_VIC_allele1", "BC18_VIC_allele2", "BC19_VIC_allele1", "BC19_VIC_allele2", 
    "BC20_VIC_allele1", "BC20_VIC_allele2", "BC24_NED_allele1", "BC24_NED_allele2", 
    "BC26_NED_allele1", "BC26_NED_allele2", "BC29_NED_allele1", "BC29_NED_allele2", 
    "BC32_PET_allele1", "BC32_PET_allele2", "BC34_PET_allele1", "BC34_PET_allele2", 
    "BC35_PET_allele1", "BC35_PET_allele2", "BC36_PET_allele1", "BC36_PET_allele2", 
    "life_stage", "YEAR", "MOUNTAIN", "wild_captive", "from_sheet")

# Other columns. Some dataframes have these columns
other_columns <- c("SOURCE_POPULATION")

# Make sure all dataframes have all common_columns
stopifnot(sapply(all_sheets, function(x) {common_columns %in% colnames(x)}))

# Get columns that aren't in common_columns or other_columns
bad_columns <- lapply(all_sheets, function(x) {
  ok_cols <- (colnames(x) %in% common_columns) | (colnames(x) %in% other_columns)
  return(colnames(x)[! ok_cols])
})

# Make sure bad columns only have NA values so we can safely remove them later
for (sheet in names(bad_columns)) {
  for (col in bad_columns[[sheet]]) {
    stopifnot(is.na(all_sheets[[sheet]][,col]))
  }
}

# Only keep good columns. Mutate everything into characters. Join everything
for (sheet in names(all_sheets)) {
  good_columns <- c(common_columns, 
                    other_columns[other_columns %in% colnames(all_sheets[[sheet]])])
  df <- all_sheets[[sheet]][,good_columns] %>% mutate_all(as.character)
  
  if (sheet == names(all_sheets)[1]) {
    all_data <- df
  } else {
    all_data <- full_join(all_data, df)
  }
}

# Remove NA rows and convert to a dataframe
genotype_data <- all_data %>% filter(! is.na(ID)) %>% as.data.frame

# Rename non-microsat columns
col_index <- ! str_detect(colnames(genotype_data), "allele[12]")
colnames(genotype_data)[col_index] <- colnames(genotype_data)[col_index] %>% 
  tolower() %>% str_replace_all("[ /]", "_")
```

```{r}
dim(genotype_data)
genotype_data %>% select(-contains("allele")) %>% head
```

------

# Fix date and IDs

All IDs should be four digits. Add a leading zero if not.
B20743 is an exception since we only have the Healsville ID. 99999 was an 
individual that escaped (new untagged individual but didn't manage to tag it
on that trapping day).

Some values in the ID column have two IDs in ####/#### format.
In this case, the last value is the current tag and the first value
should be noted as the old tag.

### Capture data

For capture data, date was encoded as days from 1900.

```{r}
# Date column encoded as days from 1900
origin_date <- "1899-12-30"
new_dates <- lubridate::days(data$date) + lubridate::ymd(origin_date)
data$date <- new_dates

data %>% head
```

Expect IDs to be 4 digits. If not, remove data and store elsewhere.

```{r}
# Some IDs are NA. Change to string.
data <- data %>% mutate(id=ifelse(is.na(id), "NA", id))

# IDs with replaced tags
# Assume second tag is the most recent one and should be used in the ID column
index <- data$id %>% str_detect("^[0-9]+/[0-9]+$") %>% which
replaced_tags <- data[index,"id"]
new_tag <- data[index,"id"] %>% str_split("/") %>% sapply(function(x) x[2])
old_tag <- data[index,"id"] %>% str_split("/") %>% sapply(function(x) x[1])
comment <- paste("Replaced tag", old_tag)

# Add old tag in the "other" column
data[index, "other"] <- append_note(comment, data[index, "other"])

# Replace ID field with new tag
data[index, "id"] <- new_tag

# View
data[index,]
```

```{r}
# Remove IDs which have non-numeric characters
not_ok <- data$id %>% str_detect("[^0-9]")
removed_data <- data[not_ok,]
data <- data[!not_ok,]

# Add leading zeroes
needs_leading_zeroes <- nchar(data$id) < 4
data$id[needs_leading_zeroes] <- 
  sprintf("%04d", as.numeric(data$id[needs_leading_zeroes]))
```

```{r}
# Fill in fields for capture data
data$year <- lubridate::year(data$date)
data$mountain <- "Mount Buller"
data$wild_captive <- "wild"
data$source_population <- NA
```

### Genotype data

```{r}
# Replace 99999 ID with untagged
genotype_data$id[genotype_data$id == "99999"] <- "untagged"

## 8651/8175 should be flipped
#genotype_data$id[genotype_data$id == "8651/8175"] <- "8175/8651"

# Samples with multiple tags
not_ok <- str_detect(genotype_data$id, "[^0-9]") & 
  (genotype_data$id != "B20743") & (genotype_data$id != "untagged")
genotype_data$id[not_ok]

# IDs with replaced tags
# Assume second tag is the most recent one and should be used in the ID column
replaced_tags <- c(replaced_tags, genotype_data[not_ok, "id"])
new_tag <- genotype_data[not_ok,"id"] %>% str_split("/") %>% sapply(function(x) x[2])
old_tag <- genotype_data[not_ok,"id"] %>% str_split("/") %>% sapply(function(x) x[1])
comment <- paste("Replaced tag", old_tag)

# Add old tag in the "other" column
genotype_data[not_ok, "notes"] <- append_note(comment, genotype_data[not_ok, "notes"])

# Replace ID field with new tag
genotype_data[not_ok, "id"] <- new_tag

# Add leading zeroes
needs_leading_zeroes <- nchar(genotype_data$id) < 4
genotype_data$id[needs_leading_zeroes] <- 
  sprintf("%04d", as.numeric(genotype_data$id[needs_leading_zeroes]))
```

### Replaced tags

```{r}
# Parse out replaced tags in genotype data notes
# genotype_data %>% select(id, retag, notes) %>% filter(retag != 0 | !is.na(notes))
txt <- genotype_data %>% select(id, date, retag, notes) %>% 
  filter(notes != "Tagged as new adult in Jan-Feb 2015") %>% 
  filter(str_detect(notes, "tag|[0-9]{4}")) %>%
  mutate(x=paste(id, notes, sep=" - ")) %>% .$x
txt

# Extract all sequences of 4 digits
replaced_tags <- txt %>% str_extract_all("[0-9]{4}") %>% lapply(unique)

# Remove replicates
replaced_tags <- replaced_tags %>% lapply(sort) %>% unique

# 2012, 2013 and 2014 refer to the year and aren't IDs
replaced_tags <- replaced_tags %>% lapply(function(x) {x[! x %in% seq(2012,2014)]})

# 2172 looks like a typo. Looks like its supposed to be 8172. Remove it.
replaced_tags <- replaced_tags %>% lapply(function(x) {x[x != "2172"]})

# Length needs to be at least 2 to be useful
replaced_tags <- replaced_tags[sapply(replaced_tags, length) > 1] %>% unique
```


```{r}
# Parse out replaced tags in capture data notes
# data %>% select(id, other) %>% filter(!is.na(other))
txt <- data %>% select(id, date, other) %>% 
  filter(str_detect(other, "tag") & str_detect(other, "[0-9]{4}")) %>%
  mutate(x=paste(id, other, sep=" - ")) %>% .$x
txt

# Extract all sequences of 4 digits
capture_replaced_tags <- txt %>% str_extract_all("[0-9]{4}") %>% lapply(unique)

# Remove replicates
capture_replaced_tags <- capture_replaced_tags %>% lapply(sort) %>% unique

# Join with previous list and remove duplicates
replaced_tags <- c(replaced_tags, capture_replaced_tags) %>% unique
```

```{r}
# 8853 and 8854 looks like a typo from the original comment:
# "Ripped ear tag from L ear. Replaced with 8853 in R ear and genetics sample taken M11BE"
stopifnot(replaced_tags[[7]] == c("8853", "8854"))
replaced_tags[[7]] <- NULL
```

```{r}
replaced_tags
```

Go through the tags and mark the most recently seen one.

```{r}
newest_replaced_tag <- rep(NA, length(replaced_tags))
for (i in seq_along(replaced_tags)) {
  newest_replaced_tag[i] <- data %>% filter(id %in% replaced_tags[[i]]) %>% 
    arrange(desc(date)) %>% .$id %>% .[1]
}
names(replaced_tags) <- newest_replaced_tag
```

```{r}
replaced_tags
```

Find any instances of old tags and replace with new tag.

```{r}
for (i in seq_along(replaced_tags)) {
  newest <- names(replaced_tags)[i]
  old <- replaced_tags[[i]][replaced_tags[[i]] != newest]
  if (any(old %in% data$id)) {
    message("Updating ", newest, " in capture data")
    data <- data %>% mutate(id=ifelse(id %in% old, newest, id))
  }
  if (any(old %in% genotype_data$id)) {
    message("Updating ", newest, " in genotype data")
    genotype_data <- genotype_data %>% mutate(id=ifelse(id %in% old, newest, id))
  }
}
replaced_tags
```

-----

# Arrange sheets

### Different samples

Check which samples in genotype data are not in capture data. All wild caputre
records in the genotype data should be in the capture data.

```{r}
genotype_id_date <- paste(genotype_data$id, genotype_data$date, sep="_")
capture_id_date <- paste(data$id, data$date, sep="_")

# How many samples from genotype data in capture data?
table(genotype_id_date %in% capture_id_date)
```

For matching individuals, sanity check that site, sex, and age match.

```{r}
# Get common samples
common <- intersect(genotype_id_date, capture_id_date)
genotype_common <- genotype_data[match(common, genotype_id_date),] %>%
   select(-contains("allele"))
capture_common <- data[match(common, capture_id_date),]
genotype_common[is.na(genotype_common)] <- ""
capture_common[is.na(capture_common)] <- ""

x <- c("site", "sex", "age", "site", "line", "point")
table(genotype_common[, x] == capture_common[, x], useNA="always")
```

Head, body, tail measurements are different though. Some are due to rounding
and some due to floating point errors. Some of the values from the genotype
data look like it's been averaged from the capture data (e.g ID 8446).

```{r}
x <- c("head", "body", "tail")
sapply(x, function(y) {table(genotype_common[,y] == capture_common[,y])})

compare <- cbind(genotype_common[,c("id",x)], capture_common[,x])
colnames(compare) <- "id head.g body.g tail.g head.c body.c tail.c" %>%
  str_split(" ") %>% unlist

table(compare$head.g == compare$head.c)
table(compare$body.g == compare$body.c)
table(compare$tail.g == compare$tail.c)

diff <- compare$head.g != compare$head.c
compare[diff,] %>% head(10)
```

Use the measurement values from the capture data.

```{r}
# Where are the individuals that aren't in the capture data from?
genotype_data[! genotype_id_date %in% capture_id_date,"from_sheet"] %>% table
```

Check if individuals from "2011 juv" sheet exist in capture data despite not
having a capture date.

```{r}
juv_2011 <- genotype_data[! genotype_id_date %in% capture_id_date,] %>% 
  filter(from_sheet == "2011 juv")
data_2011 <- data %>% filter(lubridate::year(date) == 2011) %>% 
  filter(age != 1, id %in% juv_2011$id)
table(juv_2011$id %in% data_2011$id)
data_2011$id %>% table
```

Check if individuals from "2012 juv" sheet exist in capture data despite not
having a capture date.

```{r}
juv_2012 <- genotype_data[! genotype_id_date %in% capture_id_date,] %>% 
  filter(from_sheet == "2012 juv")
data_2012 <- data %>% filter(lubridate::year(date) == 2012)
table(juv_2012$id %in% data_2012$id)
```

The individuals have records in the capture data spreadsheet, but not a one-to-one
mapping. 2012 juv genotype data are not all marked as juvenile in the capture
data. Assume that the capture spreadsheet is more acurate for juv samples age status.

What about the records that have dates, but don't match to the capture spreadsheet?

```{r}
no_match <- genotype_data[! genotype_id_date %in% capture_id_date,] %>% 
  filter(from_sheet %in% c("2011", "2013 juv", "2014")) %>%
           select(-contains("allele"))

no_match
```

```{r}
no_match %>% filter(id == "0269")
data %>% filter(id == "0269", lubridate::year(date) == 2011)
```

```{r}
no_match %>% filter(id == "0388")
data %>% filter(id == "0388", lubridate::year(date) == 2013)
```


```{r}
no_match %>% filter(id == "untagged")
data %>% filter(date == "2014-11-08", day == 4, line == 4, point == 13)
```

These three samples aren't seen in the capture data. Append these three
instances to capture records.

```{r}
add_records <- no_match %>% rename(rep_con=repro_cond, new=new_re, other=notes) %>%
  mutate(other=append_note("From genotype spreadsheet", other)) %>%
  mutate(first_weight=NA) %>% .[,colnames(data)]

data <- rbind(data, add_records)
```

### Fix notes

Some notes/comments are different between the two sheets. Move genotype data
notes to capture data notes.

```{r}
# Set NA values to empty strings for notes
genotype_data <- genotype_data %>% mutate(notes=ifelse(is.na(notes), "", notes))
data <- data %>% mutate(other=ifelse(is.na(other), "", other))

diff <- genotype_data[match(common, genotype_id_date), c("notes")] != 
  data[match(common, capture_id_date), c("other")]

change <- diff & (genotype_data[match(common, genotype_id_date), c("notes")] != "")
message("Different: ", sum(diff), "\nChanged: ", sum(change))

# Append genotype comment to existing comment
data[match(common, capture_id_date), "other"][change] <- 
  append_note(genotype_data[match(common, genotype_id_date), "notes"][change],
              data[match(common, capture_id_date), "other"][change])
```

### Fix reproduction condition

Some repoduction condition values are different between the two sheets. Move 
genotype data condition notes to capture data condition notes.

```{r}
# Set NA values to empty strings for notes
genotype_data <- genotype_data %>% 
  mutate(repro_cond=ifelse(is.na(repro_cond), "", repro_cond))
data <- data %>% mutate(rep_con=ifelse(is.na(rep_con), "", rep_con))

diff <- genotype_data[match(common, genotype_id_date), c("repro_cond")] != 
  data[match(common, capture_id_date), c("rep_con")]

change <- diff & (genotype_data[match(common, genotype_id_date), "repro_cond"] != "")
message("Different: ", sum(diff), "\nChanged: ", sum(change))

# Append rep_con comment to existing comment
data[match(common, capture_id_date), "rep_con"][change] <- 
  append_note(genotype_data[match(common, genotype_id_date), "repro_cond"][change],
              data[match(common, capture_id_date), "rep_con"][change])
```

### Isolate genotype data

Some samples have multiple genotype records with different values. Keep both.

```{r}
microsat_data <- genotype_data %>% 
  select(id, contains("allele"), -hybrid_alleles)

# Remove rows where there's no genotype data
no_data <- microsat_data %>% select(-id) %>% is.na %>% rowMeans %>% 
  magrittr::equals(1)
microsat_data <- microsat_data[! no_data,]

# Remove identical rows
microsat_data <- microsat_data[! duplicated(microsat_data),] %>%
  arrange(id)

# Remove the record that has "No Genotype" as a value in BP1_FAM_allele1
microsat_data <- microsat_data %>% filter(BP1_FAM_allele1 != "No Genotype")

# Duplicated IDs with different values
microsat_data$id %>% duplicated %>% table
dup_ids <- microsat_data$id %>% duplicated %>% which %>% 
  microsat_data$id[.] %>% unique

# Add index
microsat_data <- microsat_data %>% 
  mutate(index=seq_len(nrow(microsat_data))) %>%
  select(index, everything())

# for (samp in dup_ids) {
#   x <- microsat_data %>% filter(id == samp)
#   diff <- sapply(seq_along(x), function(y) all(! duplicated(x[,y])))
#   x[,diff]
#   break
# }

microsat_data[1:6,1:6]
```

### Non-capture records

Records which don't appear in the capture data (e.g. Hotham males 2010 & 2011, 
Timms Spur males, Zoo releases).

```{r}
non_capture <- genotype_data %>% 
  filter(from_sheet %in% c("Hotham males 2010", "Hotham males 2011",
                           "Timms Spur males 2014", "Zoo release 2013")) %>%
  select(-contains("allele"))

# Check all non-caputre records are adults
stopifnot(data$life_stage == "post-juvenile adult")

# Remove columns which are empty
empty <- non_capture %>% is.na %>% colMeans %>% magrittr::equals(1)
non_capture <- non_capture[,! empty]

# Join from sheet text to notes column and remove unnecessary columns
non_capture <- non_capture %>% 
  mutate(other=append_note(notes, from_sheet)) %>%
  select(-notes, -from_sheet, -new_re, -life_stage) %>%
  rename(rep_con=repro_cond)

# Check there aren't duplicated ids
stopifnot(! duplicated(non_capture$id))

# Join non-capture records to capture data
data <- full_join(data, non_capture)
```

-----

# Check individuals

```{r}
# All records with the same id should be the same sex
data %>% group_by(id) %>% 
  filter(sex %in% 1:2) %>%
  summarise(n_males=sum(sex==2), n=n(), percent_male=n_males/n) %>%
  filter(! percent_male %in% c(0,1))

# Manually check:
# 0268, 0398, 8054 looks like it's male
# 0269, 7140 looks like it's female
# 0594 looks like two separate animals. Female in 2002-2004, male in 2014.
# 0707 looks like two separate animals. Female in 2001, male in 2014.
# Can't tell: 0384 (1 of each and no reproduction info)

# Manually change
index <- data$id %in% c("0268", "0398", "8054")
data$sex[index] <- "2"

index <- data$id %in% c("0269", "7140")
data$sex[index] <- "1"

index <- data$id %in% c("0384")
data$sex[index] <- "not determined"
```

Some cases where there are multiple animals for a single ID.

```{r}
# Find cases where there's > 5 year span between capture dates
data %>% group_by(id) %>%
  summarise(min=min(date), max=max(date), sex=mean(as.numeric(sex))) %>%
  mutate(diff=max-min) %>%
  filter(diff > lubridate::duration(5, units="years")) %>%
  arrange(desc(diff))
```

Plot records per time.

```{r}
ggplot(data, aes(x=date, y=1)) + 
  geom_jitter(height=1, width=0, size=1, alpha=0.5)
```

```{r}
ggplot(data %>% filter(id %in% c("0379", "0707", "0594")), 
       aes(x=date, y=1, color=id)) + 
  geom_jitter(height=1, width=0, size=1, alpha=0.5)
```


```{r}
# Change 0379, 0707, 0594
# For any records before 2005, rename the sample "####_old"
index <- data$date < lubridate::ymd("2005-01-01") &
  data$id %in% c("0379", "0707", "0594")
table(index)

data[index,"id"] <- paste0(data[index,"id"], "_old")
```

Manually check possum records.

```{r}
data %>% group_by(id) %>%
  summarise(min=min(date), max=max(date), sex=mean(as.numeric(sex))) %>%
  mutate(diff=max-min) %>%
  filter(diff > lubridate::duration(4, units="years")) %>%
  arrange(desc(diff))
```

```{r}
# 7051: Assume same possum
# data %>% filter(id == "7051") %>% View

# 7728: Assume different possums before and after 2010
# data %>% filter(id == "7728") %>% View

# 8738: Same possum
# data %>% filter(id == "8738") %>% View

# 0241: Same possum
# data %>% filter(id == "0241") %>% View

# 8654: Same possum
# data %>% filter(id == "8654") %>% View

# 8470: Same
# data %>% filter(id == "8470") %>% View

# 7081: Same
# data %>% filter(id == "7081") %>% View

# 0269: Same
# data %>% filter(id == "0269") %>% View
```

```{r}
# Change 7728 pre-2010 to *_old
index <- data$date < lubridate::ymd("2010-01-01") &
  data$id %in% c("7728")
sum(index)

data[index,"id"] <- paste0(data[index,"id"], "_old")
```

-----

# Possum phenotypes

## New fields

Constant values for all possums.

```{r}
data$taxon_id <- "38600"
data$phylum <- "Chordata"
data$class <- "Mammalia"
data$order <- "Diprodontia"
data$family <- "Burramyidae"
data$genus <- "Burramys"
data$species <- "Burramys parvus"
data$subspecies <- NA
data$common_name <- "Mountain pygmy possum"
data$country <- "Australia"
data$state_or_region <- "Victoria"
```

Unknown values.

```{r}
# Don't have these values
# data$habitat <- NA
# data$decimal_latitude <- NA
# data$decimal_longitude <- NA

# Not sure if non-capture records are also by Dean Heinze. Set them as NA.
data <- data %>% 
  mutate(identified_by=ifelse(is.na(date), NA, "Dean Heinze"),
         collector=ifelse(is.na(date), NA, "Dean Heinze"))

# TODO: Do these refer to sample collection? If so, should store in another table
# data$collection_method <- NA
# data$collection_date <- NA
```

## Site

```{r}
# Convert SITE column to site_name, site_code, site_notes which can later 
# easily be combined into a 'location_text' field to comply with the OMG sample 
# metadata template.
# NB. Currently SITE values are only included for Mount Buller samples.

# Code lookup table
buller_site_lookup <- c("Federation", "Federation", "Wombat Valley", 
                       "Wombat Valley", "Women's Run", "Federation")
names(buller_site_lookup) <- c("324", "329", "331", "332", "334", "350")

# Extract numbers from SITE
site_code <- str_extract(data$site, pattern="[0-9]+")
site_names <- buller_site_lookup[site_code] %>% setNames(NULL)

# site_notes <- str_extract_all(data$site, pattern="[a-zA-Z]+")
# site_notes_full <- all_data$SITE[which(lapply(site_notes, length)>0)]

# Set location in dataframe
data$location <- paste(data$mountain, site_names, data$site, sep=", ")

# Remove site column
data <- data %>% select(-site)
```

## Sex

Change `1`s to female and `2`s to male.

```{r}
data <- data %>% 
  mutate(sex=as.character(fct_recode(as.factor(sex), female="1", male="2")))
data$sex[is.na(data$sex)] <- "not determined"

table(data$sex, useNA = "always")
```

```{r}
# If there's an obvious consensus for "not determined" individuals, use it
na_sex <- data %>% filter(sex == "not determined") %>% .$id
na_sex

# 0384 can't be determined
data %>% filter(id == "0384")

# 1420 looks female
data %>% filter(id == "1420") %>% .$sex %>% table
data <- data %>% mutate(sex=ifelse(id == "1420", "female", sex))

# 8847 looks female
data %>% filter(id == "8847") %>% .$sex %>% table
data <- data %>% mutate(sex=ifelse(id == "8847", "female", sex))

# How many in each group?
table(data$sex, useNA = "always")
```


## Body measurements

Body, head, tail, and weight

```{r}
non_numeric_values <- function(x) {
  return(x[str_detect(x, "[^0-9.]") & ! is.na(x)])
}

# Print out non-numeric values that need to be fixed
non_numeric_values(data$body)
non_numeric_values(data$head)
non_numeric_values(data$tail)
non_numeric_values(data$weight)
```

```{r}
# Replace non-numeric tail data as NA
data$tail[data$tail %in% non_numeric_values(data$tail)] <- NA

# Remove ~ and ? values in weight values
data <- data %>% mutate(weight=str_replace_all(weight, "[?~]", ""))

# Replace "nr" values as NA
data$weight[data$weight == "nr"] <- NA
```

Convert values to numeric.

```{r}
data <- data %>% mutate(
  body=as.numeric(body),
  head=as.numeric(head),
  tail=as.numeric(tail),
  weight=as.numeric(weight)
)
```

### Sanity check values

There are some values that look like errors.

```{r}
probs <- c(0, 0.01, 0.05, 0.1, 0.2,
           seq(0.25,0.75,by=0.25), 
           0.8, 0.9, 0.95, 0.98, 0.99, 0.995, 1)
apply(data[,c("body", "head", "tail", "weight")], 2, function(x) {
  quantile(x, probs=probs, na.rm=TRUE)})
```


```{r}
# Possums with weight > 70
data %>% filter(weight > 70) %>% arrange(weight) %>% 
  select(id, weight, tail:head)
```

Going to assume that anything over 100 is an error and set it to NA.

```{r}
data <- data %>% mutate(weight=ifelse(weight > 100, NA, weight))
```

Fix any obvious inconsistancies for weight over 70. Manually check other
records and set the large value as NA if it looks abnormal.

```{r}
data %>% filter(id == "0386") %>% .$weight
data %>% filter(id == "8220") %>% .$weight
data %>% filter(id == "7331") %>% .$weight
data %>% filter(id == "8175") %>% .$weight

# Set all 4 large values as NA
data <- data %>% mutate(weight=ifelse(weight > 70, NA, weight))
```


Sanity check that body should be larger than head.

```{r}
# Cases where head is larger than body
cases <- which(data$head > data$body)
non_cases <- which(data$head < data$body)
# These cases are all in Jan/Feb 2013
# data[cases,]

# How many cases?
cases %>% length
```

```{r}
# What kind of values are expected in each
data[cases,"body"] %>% summary
data[non_cases,"body"] %>% summary

data[cases,"head"] %>% summary
data[non_cases,"head"] %>% summary

# Pretty sure that cases where the recorded head size is larger than the
# recorded body size should be flipped
swap <- data$body[cases]
data$body[cases] <- data$head[cases]
data$head[cases] <- swap
```

Check some cases of large head size and small tail size in 2015.

```{r}
# data %>% filter(head > 40 & tail < 40) %>% select(id, date, weight, tail:head)
cases <- which(data$head > 40 & data$tail < 40)
length(cases)

# Summary stats
data[cases, "head"] %>% summary
data[cases, "body"] %>% summary
data[cases, "tail"] %>% summary

# Head should be body
# Body should be tail
# Tail should be head

# Swap values
swap <- data$body[cases]
data$body[cases] <- data$head[cases]
data$head[cases] <- data$tail[cases]
data$tail[cases] <- swap
```


Still one outlier with head size of 62mm

```{r}
index <- which(data$head > 50)
data[index,] %>% select(id, tail:head)

# Do other captures of the same ID have measurements? No.
data %>% filter(id == data[index, "id"]) %>% select(id, tail:head)

# Just set head size to NA
data[index, "head"] <- NA
```


There are some large males from Hotham 2011 just over 100mm, but the two females
with body size of 176mm look incorrect.

```{r}
index <- which(data$body > 105)
data[index,] %>% select(id, tail:head)

# Can we get the information from other records?.
# No other records for 0661. For 0967, future record says body size is 72.
data %>% filter(id %in% data[index, "id"]) %>% select(id, date, tail:head)

# Just set these cases to NA
data[index, "body"] <- NA
```


```{r}
# Re-check distribution of values
apply(data[,c("body", "head", "tail", "weight")], 2, function(x) {
  quantile(x, probs=probs, na.rm=TRUE)})
```

Plot histograms for each measurement.

```{r}
ggplot(data, aes(x=weight, fill=age)) +
  geom_histogram()

ggplot(data, aes(x=body, fill=age)) +
  geom_histogram()

ggplot(data, aes(x=head, fill=age)) +
  geom_histogram()

ggplot(data, aes(x=tail, fill=age)) +
  geom_histogram()
```


## Age

- age 1: adult
- age 2: juvenile
- age 3: immature adult, young adult, sub-adult

```{r}
# Two instances of non-valid age codes, id 8852. Marked as juvenile and
# adult, but one day apart.
data %>% filter(! age %in% 1:3 & ! is.na(age))

# From their rep_cond, assume 8852 is an adult
data <- data %>% mutate(age=ifelse(age %in% c("Adult", "Juvenile"), 1, age))

# How many in each group
data$age %>% table(useNA="always")
```

Check weight distribution of different ages.

```{r}
ggplot(data, aes(x=weight)) +
  geom_density() +
  facet_grid(age~.)
```

Some adults are marked as having immature juvenile pouches. These should be
age 3.

```{r}
# How many adults need to be changed?
data %>% filter(str_detect(rep_con, "juvenile"), age == "1") %>% 
  .$rep_con %>% table

data <- data %>% mutate(
  age=ifelse(rep_con == "undeveloped juvenile pouch, not pregnant", 3, age))
```

What rep conditions do juveniles have?

```{r}
data %>% filter(age == 2) %>% .$rep_con %>% table
```

What rep conditions do immature adults have?

```{r}
data %>% filter(age == 3) %>% .$rep_con %>% table
```

Recode age number as life-stage.

```{r}
data <- data %>%
  mutate(life_stage=ifelse(age == 1, "adult", NA),
         life_stage=ifelse(age == 2, "juvenile", life_stage),
         life_stage=ifelse(age == 3, "immature adult", life_stage))
```

```{r}
data$life_stage %>% table(useNA="always")
```

## Pouch young

Should contain either a number, NA, or undetermined.

```{r}
# Copy py column to number_of_pouch_young and rename py to py_raw
data <- data %>% mutate(number_of_pouch_young=py)
```

Check male values. Should be all NA.

```{r}
data %>% filter(sex == "male") %>% .$number_of_pouch_young %>% table(useNA="always")

data %>% filter(number_of_pouch_young %in% c("juv", "t10mm")) %>% 
  select(date, id:rep_con)
```

Remove juv values and assume `t10mm` refers to testis size and move to rep con.
Then set all male `number_of_pouch_young` to `NA`.

```{r}
# Replace "t10mm" value
data$rep_con[data$number_of_pouch_young == "t10mm" &
               ! is.na(data$number_of_pouch_young)] <- "10"

# Remove male number_of_pouch_young values
data <- data %>% mutate(
  number_of_pouch_young=ifelse(sex == "male", NA, number_of_pouch_young))
```

Juvenile and immature females should have zero or `NA` pouch young.

```{r}
# Check juvenile females have either NA or 0 pouch young
stopifnot(data %>% 
            filter(sex == "female" & life_stage != "adult") %>% 
            .$number_pouch_young %>% magrittr::is_in(c(NA, "0")))
```

Adult females should have a numeric value or `undetermined`.

```{r}
# Change NA values of post-juvenile females to "undetermined"
adult_females_na <- with(data, sex == "female" & 
                           life_stage == "adult" &
                           is.na(number_of_pouch_young))
adult_females_na[is.na(adult_females_na)] <- FALSE
data[adult_females_na,"number_of_pouch_young"] <- "undetermined"
```

Fix invalid values. Some values belong in rep condition.

```{r}
table(data$number_of_pouch_young)
```

```{r}
# Set `2 or 3` value to `3` since it's more likely
data <- data %>% mutate(number_of_pouch_young=ifelse(
  number_of_pouch_young == "2 or 3", "3", number_of_pouch_young))

# Set `2? Py` value to `2`
data <- data %>% mutate(number_of_pouch_young=ifelse(
  number_of_pouch_young == "2? Py", "2", number_of_pouch_young))

# Some values with Xpy and some extra rep condition information
x <- str_detect(data$number_of_pouch_young, "\\dpy") & 
  ! is.na(data$number_of_pouch_young)
data$number_of_pouch_young[x]

# Separate as number of pouch young and rep condition
match <- data$number_of_pouch_young[x] %>% str_match("^(\\d)py (\\w+)$")
match

# Replace values in pouch young column
data$number_of_pouch_young[x] <- match[,2]

# Append rep condition
data$rep_con[x] <- append_note(match[,3], data$rep_con[x])
```

For the remaining values, append to rep condition.

```{r}
data %>% filter(! number_of_pouch_young %in% c(0,1,2,3,4,"undetermined",NA)) %>%
  .$number_of_pouch_young %>% unique

# Append note
x <- ! data$number_of_pouch_young %in% c(0,1,2,3,4,"undetermined",NA)
data$rep_con[x] <- append_note(data$number_of_pouch_young[x], data$rep_con[x])
data$number_of_pouch_young[x] <- "undetermined"
```

How many in each group?

```{r}
table(data$number_of_pouch_young, useNA="always")
```

## Reproductive condition

### Move notes from Other column

```{r}
data <- data %>% mutate(misc_reproductive_notes="")
```

Some values in the "other" coumn are reproductive related

```{r}
# Manually going through, these values should be in rep_con:
# half sw nip, p stretched, no milk, likely preg
# pos preg / likely preg
# lop nip 1mm
# p not moist, preg maybe, half sw nip, p stretched, bred last year

data$other %>% str_detect("preg|\\dmm") %>% data$other[.]
```

```{r}
# Manually change the 6 entries

index <- data$other == "half sw nip, p stretched, no milk, likely preg"
data[index,"other"] <- ""
data[index,"rep_con"] <- append_note("likely preg; half sw nip, no milk",
                                     data[index,"rep_con"])  
data[index,"misc_reproductive_notes"] <- "p stretched"

index <- data$other == "pos preg"
data[index,"other"] <- ""
data[index,"misc_reproductive_notes"] <- "pos preg"

index <- data$other == "lop nip 1mm"
data[index,"other"] <- ""
data[index,"rep_con"] <- append_note("lop nip 1mm", data[index,"rep_con"])

index <- data$other == "left eye missing, mites around face, bred previously, not preg this time"
data[index,"other"] <- "left eye missing, mites around face"
data[index,"misc_reproductive_notes"] <- "bred previously, not preg this time"

index <- data$other == "p not moist, preg maybe, half sw nip, p stretched, bred last year"
data[index,"other"] <- ""
data[index,"rep_con"] <- append_note("half sw nip", data[index,"rep_con"])  
data[index,"misc_reproductive_notes"] <- "preg maybe, bred last year, p not moist, p stretched"

index <- data$other == "likely preg"
data[index,"other"] <- ""
stopifnot(data[index,"rep_con"] == "lmop; lmop like preg.")
data[index,"rep_con"] <- "lmop; likely preg"
```

```{r}
data <- data %>% mutate(reproductive_condition=rep_con)
```

Handle `2nd Litter values` by moving them to the other column.

```{r}
data %>% filter(str_detect(reproductive_condition, "2nd|second")) %>% 
  .$reproductive_condition %>% table

prob_second <- str_detect(data$reproductive_condition, "2nd|second")

# Remove from reproductive condition
data[prob_second, "reproductive_condition"] <- 
  data[prob_second, "reproductive_condition"] %>%
  str_replace("(; )?(2nd Litter\\?)|(, prob 2nd litter)|(, second litter)", "") %>%
  str_replace("second or late litter", "")

# Add "2nd litter?" to notes column
data[prob_second, "misc_reproductive_notes"] <- 
  append_note("2nd litter?", data[prob_second, "misc_reproductive_notes"])
```

```{r}
# Remove unnecessary value
index <- data$reproductive_condition == "sex wasn't recorded"
data[index,"reproductive_condition"] <- ""
```

### Male conditions

#### Testes condition

Handle male reproductive info. This refers to testis size conditions.

From email: 5mm testes or less - these are normal size testes (fully contracted)

(Remove `testis_size_status` for now. Until there's an easy threshold size to
classify size into normal/enlarge.)

```{r}
# Set testis_size_status as either "normal" or "enlarged" for adult males, 
# and NA for females and juvenile males. Enlarged males are identified as 
# having a numerical figure in rep_con.

# Some juveniles have values in reproductive condition
data %>% filter(sex == "male", life_stage != "adult", 
                reproductive_condition != "") %>%
  mutate(x=paste(life_stage, reproductive_condition, sep=" - ")) %>%
  .$x

# Set testis_size_status
# data <- data %>%
#   mutate(testis_size_status=ifelse(sex == "male" & life_stage == "adult",
#                                    "normal", NA)) %>%
#   mutate(testis_size_status=ifelse(sex == "male" & life_stage == "adult" &
#                                      str_detect(reproductive_condition, "\\d"),
#                                    "enlarged", testis_size_status))
```

#### Testes size

Handle cases where there's numeric data only.

```{r}
# Male possums with numerical values
testis_info <- data$sex == "male" & ! is.na(data$reproductive_condition) &
  str_detect(data$reproductive_condition, "^\\d+(.\\d)?$")

# What values?
data[testis_info,"reproductive_condition"] %>% table

# Set testis_size_in_mm
data$testis_size_in_mm <- NA
data$testis_size_in_mm[testis_info] <- data[testis_info,"reproductive_condition"]

# Remove values from repro condition
data[testis_info,"reproductive_condition"] <- ""
```

Some cases were the values are `x by y`. Use `x` value.  
Some cases were the values are `x * 2`. Use `x` value.

```{r}
# Get matching possums
testis_info <- data$sex == "male" & ! is.na(data$reproductive_condition) &
  str_detect(data$reproductive_condition, "by|\\*")

# What values?
data[testis_info,"reproductive_condition"] %>% unique

# Extract first number
match <- data[testis_info, "reproductive_condition"] %>% 
  str_match("^(\\d+(\\.\\d)?)(.+)$")
match %>% head

# Set testis_size_in_mm
data$testis_size_in_mm[testis_info] <- match[,2]

# Remove values from repro condition
data[testis_info,"reproductive_condition"] <- ""
```

Manually tidy up specific strings which are in the format of `Txmm`.

```{r}
data$reproductive_condition[data$reproductive_condition == "T6"] <- "6mm tes"
data$reproductive_condition[data$reproductive_condition == "T12mm"] <- "12mm tes"
data$reproductive_condition[data$reproductive_condition == "T6x12"] <- "12mm tes"
```

Handle cases of `xmm tes`.

```{r}
# Get matching possums
testis_info <- data$sex == "male" & ! is.na(data$reproductive_condition) &
  str_detect(data$reproductive_condition, "mm tes")

# What values?
data[testis_info,"reproductive_condition"] %>% unique

# Extract first number
match <- data[testis_info, "reproductive_condition"] %>% 
  str_replace("mm tes", "")

# Set testis_size_in_mm
data$testis_size_in_mm[testis_info] <- match

# Remove values from repro condition
data[testis_info,"reproductive_condition"] <- ""
```

Handle remaining cases manually.

Those with numerical values.

```{r}
# Values to fix containing a numerical value
x <- data$sex == "male" & ! is.na(data$reproductive_condition) & 
  str_detect(data$reproductive_condition, "\\d")
data[x,"reproductive_condition"]

# "6, no orange tinge"
index <- data$reproductive_condition == "6, no orange tinge"
data[index,"testis_size_in_mm"] <- "6"
data[index,"reproductive_condition"] <- ""
data[index,"other"] <- append_note("no orange tinge", data[index,"other"])

# "6, grey fur around genitals"
index <- data$reproductive_condition == "6, grey fur around genitals"
data[index,"testis_size_in_mm"] <- "6"
data[index,"reproductive_condition"] <- ""
data[index,"other"] <- append_note("grey fur around genitals", data[index,"other"])

# "<5"
index <- data$reproductive_condition == "<5"
data[index,"reproductive_condition"] <- ""
data[index,"misc_reproductive_notes"] <- 
  append_note("testes size < 5", data[index,"misc_reproductive_notes"])
```

Remaining values should go into the `other` or `misc_reproductive_notes` column.

```{r}
# Values to fix containing a numerical value
x <- data$sex == "male" & ! is.na(data$reproductive_condition) & 
  str_detect(data$reproductive_condition, "tes")
data[x,"reproductive_condition"]

# Append to notes in misc_reproductive_notes
data[x,"misc_reproductive_notes"] <- 
  append_note(data[x,"reproductive_condition"], data[index,"misc_reproductive_notes"])

# Remove values from repro condition
data[x,"reproductive_condition"] <- ""
```

```{r}
# Values to fix containing a numerical value
x <- data$sex == "male" & ! is.na(data$reproductive_condition) & 
  data$reproductive_condition != ""
data[x,"reproductive_condition"]

# Append to notes in other
data[x,"other"] <- append_note(data[x,"reproductive_condition"], data[index,"other"])

# Remove values from repro condition
data[x,"reproductive_condition"] <- ""
```

```{r}
# Check that all repro condition values should be blank for males
stopifnot(data %>% filter(sex == "male") %>% .$reproductive_condition %>% 
  magrittr::equals(""))
```

### Female conditions

```{r}
# data %>% select(date, id, sex,reproductive_condition, number_of_pouch_young) %>%
#   filter(sex == "female")
```

#### Pregnancy and breeding status

Create new column for pregnancy status.  
Possible values: not pregnant, likely not pregnant, likely pregnant, pregnant

```{r}
data <- data %>% mutate(pregnancy_status="")
```

Bred status (first year not pregnant and bred). 
Move to `misc_reproductive_notes`.

```{r}
# Get matching possums with pregnancy status
regex_string <- "1st year not preg|(prob )?bred in spring|bred\\?|bred last y(ear|r)|bred before|not bred|pos bred in prev years|prev had young\\?"
index <- str_detect(data$reproductive_condition, regex_string) &
  data$life_stage == "adult"

# What values?
data[index,"reproductive_condition"] %>% unique

# Match string
match <- data[index,"reproductive_condition"] %>% 
  str_match(regex_string)

# Add to misc_reproductive_notes
data[index,"misc_reproductive_notes"] <- 
  append_note(match[,1], data[index,"misc_reproductive_notes"])

# Remove bred condition in reproductive_condition
data[index,"reproductive_condition"] <- 
  data[index,"reproductive_condition"] %>% str_replace_all(regex_string, "")
```

Two cases to handle manually.

```{r}
index <- str_detect(data$reproductive_condition, "1st year|bred") &
  data$life_stage == "adult"

# What values?
data[index,"reproductive_condition"] %>% unique

# "bred ldp"
index <- data$reproductive_condition == "bred ldp"
data[index,"reproductive_condition"] <- "ldp"
data[index,"misc_reproductive_notes"] <- "bred"

# "pl?? May not have bred this year"
index <- data$reproductive_condition == "pl?? May not have bred this year"
data[index,"reproductive_condition"] <- ""
data[index,"misc_reproductive_notes"] <- "Post lactating?; May not have bred this year"
```

Likely not pregnant statuses. Handle these manually.

```{r}
# Get matching possums with pregnancy status
x <- str_detect(data$reproductive_condition, "not preg\\?|prob not preg|pos not preg")

# What values?
data[x,"reproductive_condition"] %>% unique

# "lop not preg?"
index <- data$reproductive_condition == "lop not preg?"
data[index,"reproductive_condition"] <- "lop"
data[index,"pregnancy_status"] <- "likely not pregnant"

# "lop not moist/greasy prob not preg"
# Change not moist/greasy to dry
index <- data$reproductive_condition == "lop not moist/greasy prob not preg"
data[index,"reproductive_condition"] <- "ldop"
data[index,"pregnancy_status"] <- "likely not pregnant"

# "ldop pos not preg, "
index <- data$reproductive_condition == "ldop pos not preg, "
data[index,"reproductive_condition"] <- "ldop"
data[index,"pregnancy_status"] <- "likely not pregnant"
data[index,"misc_reproductive_notes"] <- 
  append_note("pos bred in prev years", data[index,"misc_reproductive_notes"])
```

Not pregnant statuses.

```{r}
# Get matching possums with pregnancy status
index <- str_detect(data$reproductive_condition, "not preg|(n. preg)")

# What values?
data[index,"reproductive_condition"] %>% unique

# Set pregnancy status
data[index,"pregnancy_status"] <- "not pregnant"

# Remove status from reproductive_condition
data[index,"reproductive_condition"] <- 
  data[index,"reproductive_condition"] %>% 
  str_replace("not preg(nant)?|(n. preg)", "")
```

Likley pregnant cases.

```{r}
# Get matching possums with pregnancy status
regex_string <- "preg(nant)?\\?|prob preg|like(ly)? preg(\\.)?|pos (now )?preg"
index <- str_detect(data$reproductive_condition, regex_string)

# What values?
data[index,"reproductive_condition"] %>% unique

# Set pregnancy status
data[index,"pregnancy_status"] <- "likely pregnant"

# Remove status from reproductive_condition
data[index,"reproductive_condition"] <- data[index,"reproductive_condition"] %>% 
  str_replace(regex_string, "")
```

Pregnant cases.

```{r}
# Get matching possums with pregnancy status
regex_string <- "(heavily )?preg(nant)?(, birth imminent)?"
index <- str_detect(data$reproductive_condition, regex_string)

# What values?
data[index,"reproductive_condition"] %>% unique

# Set pregnancy status
data[index,"pregnancy_status"] <- "pregnant"

# Remove status from reproductive_condition
data[index,"reproductive_condition"] <- data[index,"reproductive_condition"] %>% 
  str_replace_all(regex_string, "")
```

Check there are no pregnancy values left.

```{r}
stopifnot(! str_detect(data$reproductive_condition, "preg"))
```

#### Size of young

If multiple young, take the value of the largest.

Values with all numerical values refer to pouch young size in mm.

```{r}
data <- data %>% mutate(pouch_young_size_in_mm="")
```

```{r}
# Get matching possums
index <- str_detect(data$reproductive_condition, "^\\d+$")

# What values?
data[index,"reproductive_condition"] %>% table

# Add number to pouch_young_size_in_mm
data[index, "pouch_young_size_in_mm"] <- data[index,"reproductive_condition"]

# Remove status from reproductive_condition
data[index,"reproductive_condition"] <- ""
```

Some have values in the format of `xmm`

```{r}
# Get matching possums
index <- str_detect(data$reproductive_condition, "^\\d+mm$")

# What values?
data[index,"reproductive_condition"] %>% table

# Add number to pouch_young_size_in_mm
data[index, "pouch_young_size_in_mm"] <- 
  data[index,"reproductive_condition"] %>%
  str_replace("mm$", "")

# Remove status from reproductive_condition
data[index,"reproductive_condition"] <- ""
```

Handle neonate cases manually.

```{r}
# Get matching possums
x <- str_detect(data$reproductive_condition, "extra|neonate|born")

# What values?
data[x,"reproductive_condition"] %>% unique

# "py prob born 2 nights ago"
index <- data$reproductive_condition == "py prob born 2 nights ago"
data[index,"misc_reproductive_notes"] <- 
  append_note("Pouch young probably born 2 nights ago",
              data[index,"misc_reproductive_notes"])
data[index,"reproductive_condition"] <- ""

# "6 born o/n"
index <- data$reproductive_condition == "6 born o/n"
data[index,"misc_reproductive_notes"] <- 
  append_note("6 neonates",
              data[index,"misc_reproductive_notes"])
data[index,"reproductive_condition"] <- ""

# "6 plus 1 neonate born overnight"
index <- data$reproductive_condition %in% 
  c("6 plus 1 neonate born overnight", "6mm plus 1 born overnight")
data[index,"pouch_young_size_in_mm"] <- "6"
data[index,"misc_reproductive_notes"] <- 
  append_note("1 extra neonate", data[index,"misc_reproductive_notes"])
data[index,"reproductive_condition"] <- ""

# "6 plus 4 neonates born overnight"
index <- data$reproductive_condition == "6 plus 4 neonates born overnight"
data[index,"pouch_young_size_in_mm"] <- "6"
data[index,"misc_reproductive_notes"] <- 
  append_note("4 extra neonates", data[index,"misc_reproductive_notes"])
data[index,"reproductive_condition"] <- ""

# "6 plus 2 neonates"
index <- data$reproductive_condition == "6 plus 2 neonates"
data[index,"pouch_young_size_in_mm"] <- "6"
data[index,"misc_reproductive_notes"] <- 
  append_note("2 extra neonates", data[index,"misc_reproductive_notes"])
data[index,"reproductive_condition"] <- ""

# "5mm, plus 1 extra young born overnight"
index <- data$reproductive_condition == "5mm, plus 1 extra young born overnight"
data[index,"pouch_young_size_in_mm"] <- "5"
data[index,"misc_reproductive_notes"] <- 
  append_note("1 extra neonate", data[index,"misc_reproductive_notes"])
data[index,"reproductive_condition"] <- ""

# "5mm, plus 2 extra young born overnight"
index <- data$reproductive_condition == "5mm, plus 2 extra young born overnight"
data[index,"pouch_young_size_in_mm"] <- "5"
data[index,"misc_reproductive_notes"] <- 
  append_note("2 extra neonates", data[index,"misc_reproductive_notes"])
data[index,"reproductive_condition"] <- ""

# "6 plus 5 neonates"
index <- data$reproductive_condition == "6 plus 5 neonates"
data[index,"pouch_young_size_in_mm"] <- "6"
data[index,"misc_reproductive_notes"] <- 
  append_note("5 extra neonates", data[index,"misc_reproductive_notes"])
data[index,"reproductive_condition"] <- ""

# "5mm probably born overnight"
index <- data$reproductive_condition == "5mm probably born overnight"
data[index,"pouch_young_size_in_mm"] <- "5"
data[index,"misc_reproductive_notes"] <- 
  append_note("Probaby gave birth overnight", data[index,"misc_reproductive_notes"])
data[index,"reproductive_condition"] <- ""

stopifnot(data[x,"reproductive_condition"] == "")
```

A few cases with multiple young to handle manually.

```{r}
### save point
# save.image(file="cache/image_2017-11-02.RData")
# load("cache/image_2017-11-02.RData")
```


```{r}
# Get matching possums
x <- str_detect(data$reproductive_condition, "^\\d{2}|20mm") &
  ! str_detect(data$reproductive_condition, ";")

# What values?
data[x,"reproductive_condition"] %>% unique

# "146?" should be a comment
index <- data$reproductive_condition == "146?"
data[index,"other"] <- append_note("146?", data[index,"other"])
data[index,"reproductive_condition"] <- ""

# "25 grey"
index <- data$reproductive_condition == "25 grey"
data[index,"pouch_young_size_in_mm"] <- "25"
data[index,"reproductive_condition"] <- ""

# "4PY 20mm"
index <- data$reproductive_condition == "4PY 20mm"
data[index,"number_of_pouch_young"] <- "4"
data[index,"pouch_young_size_in_mm"] <- "20"
data[index,"reproductive_condition"] <- ""
```

Some values have two numbers. Take the largest one.

```{r}
# Get matching possums
x <- str_detect(data$reproductive_condition, "\\d; \\d")

# What values?
data[x,"reproductive_condition"] %>% unique

# Get largest pouch young
largest_size <- data[x,"reproductive_condition"] %>% 
  str_split("; ") %>% 
  sapply(function(sizes) {max(as.numeric(sizes))})

# Get smallest pouch young
smallest_comment <- data[x,"reproductive_condition"] %>% 
  str_split("; ") %>% 
  sapply(function(sizes) {min(as.numeric(sizes))}) %>%
  paste0("Additional pouch young of size ", ., "mm")

# Insert values
data[x,"pouch_young_size_in_mm"] <- largest_size
data[x,"misc_reproductive_notes"] <- 
  append_note(smallest_comment, data[x,"misc_reproductive_notes"])

# Remove values
data[x, "reproductive_condition"] <- ""
```

#### Pouch condition

```{r}
data <- data %>% mutate(pouch_condition="")
```

Handle cases of juvenile pouches.

```{r}
# Which possums are marked as adult, but rep condition indicates juvenile
discordant <- data %>% filter(life_stage == "adult",
                str_detect(reproductive_condition, "[Jj]uv")) %>%
  select(reproductive_condition, life_stage) %>% unique
discordant

# These should be immature adults
regex_string <- "juv p(ouch)?|Juv P LP unused clean"
x <- str_detect(data$reproductive_condition, regex_string)
data[x, "reproductive_condition"] <- data[x, "reproductive_condition"] %>% 
  str_replace(regex_string, "")
data[x,"life_stage"] <- "immature adult"

# Other juvenile values
x <- str_detect(data$reproductive_condition, "[Jj]uv")
stopifnot(data[x, "life_stage"] != "adult")

# What values?
data[x,"reproductive_condition"] %>% unique

# Remove values since they're uninformative
data[x,"reproductive_condition"] <- ""
```

Juvenile reproductive conditions.

```{r}
# There's one instance (id 1932, date 18/2/15, weight 29) of a juvenile 
# marked as post-lactating. Assume that the life stage is accurate and 
# reproductive condition is incorrect.
# data %>% filter(life_stage == "juvenile", reproductive_condition != "") %>%
#   select(id, date, life_stage, reproductive_condition)

# Values
x <- ! is.na(data$life_stage) & data$life_stage == "juvenile" & 
  data$reproductive_condition != ""
data[x,"reproductive_condition"] %>% unique

# Remove unneeded information
new_value <- data[x,"reproductive_condition"] %>% 
  str_replace("Post lactating|clean pouch|nolac|nonlac|no signs( of lac)?", "")

# Remove reproductive condition
data[x,"reproductive_condition"] <- new_value
```

Handle immature adult cases.

```{r}
# Values
x <- ! is.na(data$life_stage) & data$life_stage == "immature adult" & 
  data$reproductive_condition != ""
data[x,"reproductive_condition"] %>% unique

# Clean pouch status is unncessary since they're immature adults
regex_string <- "juv pouch|undeveloped juvenile pouch, |not bred juv p|clean pouch juv cond|clean pouch|undev pouch|clean"
data[x,"reproductive_condition"] <- data[x,"reproductive_condition"] %>%
  str_replace(regex_string, "")
```

Manually handle weird cases. Sometimes "dry" refers to pouch and sometimes
nipples.

```{r}
x <- str_detect(data$reproductive_condition, "dry|clean|moist")

# What values?
data[x,"reproductive_condition"] %>% unique

find_replace <- list(
  "ldp dry ; lmdop" = "dry lmdop",
  "dry p, , , lop nip 1.5mm" = "dry lop; nip 1.5mm",
  "lop ; dry p, , , lop nip 1.5mm" = "dry lop; nip 1.5mm",
  "post lactating, fully dry contracted pouch and nipples" =
    "dry p; post lactating; contracted pouch and nipples",
  "lop dry abort" = "dry lop; abort",
  "dry, " = "dry p",
  "dry " = "dry p",
  "4 sw nip, dry, " = "4swn dry",
  "clean pouch" = "clean p",
  "lop  orange moist" = "lmop",
  "moist p" = "mp",
  "lmop , wet and moist pouch may give birth soon" = 
    "lmop; may give birth soon"
)

for (old in names(find_replace)) {
  data <- data %>% mutate(
    reproductive_condition=ifelse(reproductive_condition == old, 
                                  find_replace[[old]], reproductive_condition))
}
```

Pouch status: loose, moist, dark, orange, etc.

```{r}
regex_string <- "(\\b|^)(clean )?(dry )?(moist )?([Ll]?[Mm]?[Dd]?[Oo]?[Pp])(\\b|$)"
x <- str_detect(data$reproductive_condition, regex_string)

# What values?
data[x,"reproductive_condition"] %>% unique %>% head

# Regex match
match <- data[x,"reproductive_condition"] %>%
  str_match(regex_string)
match[,1] %>% tolower %>% unique

# Move pouch condition to separate column
data[x, "pouch_condition"] <- match[,1] %>% tolower

# Remove pouch condition from rep condition
data[x,"reproductive_condition"] <- data[x,"reproductive_condition"] %>%
  str_replace(match[,1], "")
```

Manually edit a few cases.

```{r}
# "lodp"
index <- data$reproductive_condition == "lodp"
stopifnot(data[index,"pouch_condition"] == "")
data[index,"pouch_condition"] <- "ldop"
data[index,"reproductive_condition"] <- ""
```

#### Fat in pouch

FIP status
```{r}
data$fat_in_pouch <- ""
```

```{r}
regex_string <- "((No|no|half|quarter|little|some) )?(fip|FIP)"
x <- str_detect(data$reproductive_condition, regex_string)

# What values?
data[x,"reproductive_condition"] %>% unique %>% head

new_value <- data[x, "reproductive_condition"] %>% 
  str_extract(regex_string) %>%
  tolower %>% 
  str_replace("fip", "FIP")

# Set value and remove from reproductive_condition
data[x, "fat_in_pouch"] <- new_value
data[x, "reproductive_condition"] <- data[x, "reproductive_condition"] %>% 
  str_replace(regex_string, "")

# Check no FIP notes left
stopifnot(! str_detect(data$reproductive_condition, "FIP|fip"))
```

#### Lactation status

```{r}
data <- data %>% mutate(lactation_status="")
```

Not lactating possums.

```{r}
# How many possums have non-lactating status
values <- c("Not lac", "nonlac", "no signs of lac")
sum(data$reproductive_condition %in% values)

# Replace values
data <- data %>% mutate(
  lactation_status=ifelse(reproductive_condition %in% values, 
                          "not lactating", lactation_status),
  reproductive_condition=ifelse(reproductive_condition %in% values, 
                                "", reproductive_condition)
)
```

Advanced post lactating possums.  
(These possums should be putting on weight, nipples contracted, 
approx. 1 month since having litter).

```{r}
x <- str_detect(data$reproductive_condition, "apl")

# What values?
data[x,"reproductive_condition"] %>% unique

# Move condition to separate column
data[x, "lactation_status"] <- "advanced post lactating"

# Remove lactation status from rep condition
data[x,"reproductive_condition"] <- data[x,"reproductive_condition"] %>%
  str_replace("apl", "")
```

Post lactacting possums.

```{r}
regex_string <- "pl|(recent )?[Pp]ost( [Ll](act)?(ating)?)?(, but more recent than others)?|\\dx previously lac|finished lac|(recent )?PL"
x <- str_detect(data$reproductive_condition, regex_string)

# What values?
data[x,"reproductive_condition"] %>% unique %>% head

# Regex match
match <- data[x,"reproductive_condition"] %>%
  str_match(regex_string)
match[,1] %>% tolower %>% unique

# Move condition to separate column
data[x, "lactation_status"] <- "post lactating"

# Remove lactation status from rep condition
data[x,"reproductive_condition"] <- data[x,"reproductive_condition"] %>%
  str_replace(regex_string, "")
```

Lactating possums often have nipple information as well. Parse these later.

#### Nipple condition

```{r}
data <- data %>% mutate(nipple_condition="")
```

Lactating and recently lactating.

```{r}
regex_string <- "(recently( )?)?lac(t(ating)?)?\\??"
x <- str_detect(data$reproductive_condition, regex_string)

# What values?
data[x,"reproductive_condition"] %>% unique

# Move condition to separate column
data[x, "lactation_status"] <- "lactating"

# Manual changes
data[data$reproductive_condition == "4x4lac", "reproductive_condition"] <- 
  "4x4 lac"
data[data$reproductive_condition == "lac4x4", "reproductive_condition"] <- 
  "4x4 lac"
data[data$reproductive_condition == "2x4lac", "reproductive_condition"] <- 
  "2x4 lac"
data[data$reproductive_condition == "recentlylac", "reproductive_condition"] <- 
  "recently lac"

# "4sw nip/4 lact nip/abort?"
index <- data$reproductive_condition == "4sw nip/4 lact nip/abort?"
data[index,"reproductive_condition"] <- ""
data[index,"nipple_condition"] <- "4ln; 4swn"
stopifnot(data[index,"other"] == "abort?")

# "4 sw nip, no milk, lact?"
index <- data$reproductive_condition == "4 sw nip, no milk, lact?"
data[index,"reproductive_condition"] <- ""
data[index,"nipple_condition"] <- "4swn; no milk"

# Move nipple condition to separate column
data[x,"nipple_condition"] <- data[x,"reproductive_condition"]
data[x,"reproductive_condition"] <- ""
```

Lactating nipples in the format of `4ln`.

```{r}
regex_string <- "ln"
x <- str_detect(data$reproductive_condition, regex_string)

# What values?
data[x,"reproductive_condition"] %>% unique

# Mark these as lactating
data[x, "lactation_status"] <- "lactating"

# Manulally change "4ln milk, young in nest; 4 ln milk"
index <- data$reproductive_condition == "4ln milk, young in nest; 4 ln milk"
data[index, "reproductive_condition"] <- "4ln milk"
data[index, "other"] <- append_note("young in nest", data[index, "other"])

# Manually replace values with multiple values
find_replace <- list(
  "4ln milk; 4ln milk" = "4ln milk",
  "4ln half milk; 4ln" = "4ln half milk",
  "4ln milk; 4ln" = "4ln milk",
  "4ln 3 quarter milk; 4ln" = "4ln 3 quarter milk",
  " ; 4ln" = "4ln"
)
for (old in names(find_replace)) {
  data <- data %>% mutate(
    reproductive_condition=ifelse(reproductive_condition == old, 
                                  find_replace[[old]], reproductive_condition))
}

# Move condition to separate column
data[x,"nipple_condition"] <- 
  data[x,"reproductive_condition"] %>% str_trim()

# Remove lactation status from rep condition
data[x,"reproductive_condition"] <- ""
```

Lactating nipples in the format of `4x` or `4t` or `4 teets`.

```{r}
regex_string <- "\\dx|\\dt|\\d teets"
x <- str_detect(data$reproductive_condition, regex_string)

# What values?
data[x,"reproductive_condition"] %>% unique

# Manulally change "4x enlarged teets (grumpy!)"
index <- data$reproductive_condition == "4x enlarged teets (grumpy!)"
data[index, "reproductive_condition"] <- "4swn"
data[index, "other"] <- append_note("grumpy!", data[index, "other"])

# 2 "4x " values are from possums who are post lactating
data[data$reproductive_condition == "4x ", "reproductive_condition"] <- ""

# Standardize values
new_value <- data[x,"reproductive_condition"] %>% 
  str_replace("(\\d) teets", "\\1x") %>%
  str_replace("(\\d)t", "\\1x")
new_value %>% unique

# Move condition to separate column
data[x,"nipple_condition"] <- new_value

# Remove lactation status from rep condition
data[x,"reproductive_condition"] <- ""

# Mark those with "4x milk" as lactating
data[data$nipple_condition == "4x milk", "lactation_status"] <- "lactating"
```

Contracted nipples. Discard information on contracted pouches (not may cases).

```{r}
regex_string <- "contract"
x <- str_detect(data$reproductive_condition, regex_string)

# What values?
data[x,"reproductive_condition"] %>% unique

# "contracted " values refer to pouches. Remove these values
# data %>% filter(reproductive_condition == " contracted ")
data[data$reproductive_condition == " contracted ", "reproductive_condition"] <- ""

# Clean values
new_values <- data[x,"reproductive_condition"] %>%
  str_replace("^[,; ]+", "") %>%
  str_replace(" P and N", "") %>%
  str_replace(" (nipes|nip)$", " nipples") %>%
  str_replace(" pouch/nipples", "") %>%
  str_replace(" pouch and nipples", "") %>%
  str_replace(" nipples", "")
new_values %>% unique

# Move condition to separate column
data[x,"nipple_condition"] <- new_values

# Remove lactation status from rep condition
data[x,"reproductive_condition"] <- ""
```

Swollen nipples.

```{r}
regex_string <- "sw nip"
x <- str_detect(data$reproductive_condition, regex_string)

# What values?
data[x,"reproductive_condition"] %>% unique

# Manual changes
index <- data$reproductive_condition == 
  ", no sw nip, saggy mam gland, lots orange fur, dry"
data[index,"misc_reproductive_notes"] <- append_note(
  "no sw nip, saggy mam gland", data[index,"misc_reproductive_notes"])
data[index,"reproductive_condition"] <- ""

index <- data$reproductive_condition == "4 half sw nip, dark blotches and orange fur"
data[index,"nipple_condition"] <- "4 half swn, dark blotches"
data[index,"other"] <- append_note("orange fur", data[index,"other"])
data[index,"reproductive_condition"] <- ""

# Clean values
x <- str_detect(data$reproductive_condition, regex_string)
new_values <- data[x,"reproductive_condition"] %>%
  str_replace("^[,; ]+", "") %>%
  str_replace("nipples", "nip") %>%
  str_replace("sw nip", "swn") %>% 
  str_replace("3swn", "3 swn")
new_values %>% unique

# Move condition to separate column
data[x,"nipple_condition"] <- new_values

# Remove lactation status from rep condition
data[x,"reproductive_condition"] <- ""
```

Extract size.

```{r}
regex_string <- "~|mm"
x <- str_detect(data$reproductive_condition, regex_string)

# What values?
data[x,"reproductive_condition"] %>% unique

# Clean values
new_values <- data[x,"reproductive_condition"] %>% 
  str_match("\\d(\\.\\d)?") %>% .[,1] %>% paste0(., "mm")
new_values %>% unique

# Move to nipple_condition column
data[x,"nipple_condition"] <- 
  append_note(new_values, data[x,"nipple_condition"])

# Remove lactation status from rep condition
data[x,"reproductive_condition"] <- ""
```

Other values. Handle manually.

```{r}
x <- str_detect(data$reproductive_condition, "nip|sn|swn")

# What values?
data[x,"reproductive_condition"] %>% unique

index <- data$reproductive_condition == ", orange fur, nip quarter size"
data[index,"nipple_condition"] <- "quarter size"
data[index,"other"] <- append_note("orange fur", data[index,"other"])
data[index,"reproductive_condition"] <- ""

index <- data$reproductive_condition == ", dry retracted nip, dark blotches"
data[index,"nipple_condition"] <- "dry retracted, dark blotches"
data[index,"reproductive_condition"] <- ""

index <- data$reproductive_condition == ", 2 sw. nip., skinny"
data[index,"nipple_condition"] <- "2 swn"
data[index,"other"] <- append_note("skinny", data[index,"other"])
data[index,"reproductive_condition"] <- ""

index <- data$reproductive_condition == " 4swn aborted, , moist abundant orange hair"
data[index,"nipple_condition"] <- "4 swn"
data[index,"other"] <- append_note("aborted; moist abundant orange hair",
                                   data[index,"other"])
data[index,"reproductive_condition"] <- ""

index <- data$reproductive_condition == "4swn 0 milk dry recent wean"
data[index,"nipple_condition"] <- "4 swn, dry, no milk, recent wean"
data[index,"reproductive_condition"] <- ""

index <- data$reproductive_condition == "4sn dry"
data[index,"nipple_condition"] <- "4 swn, dry"
data[index,"reproductive_condition"] <- ""

index <- data$reproductive_condition == "4swn dry"
data[index,"nipple_condition"] <- "4 swn, dry"
data[index,"reproductive_condition"] <- ""

# Check remaining values are handled
stopifnot(data[x,"reproductive_condition"] == "")
```


Manually fix instance of `4x4` which should be `4x4 finished lac`

```{r}
stopifnot(data[data$nipple_condition == "4x4","rep_con"] == "4x4finished lac")
data[data$nipple_condition == "4x4","nipple_condition"] <- "4x4 finished lac"
```

Add space between number and `ln` (e.g. `4ln` to `4 ln`).

```{r}
data$nipple_condition <- data$nipple_condition %>% 
  str_replace("(\\d)ln", "\\1 ln")
```


#### Remaining values

```{r}
x <- data$reproductive_condition != ""

# Clean up values
new_values <- data$reproductive_condition[x] %>%
  str_replace("^[ ;,.]+", "") %>%
  str_replace("[ ;,]+$", "")
new_values %>% unique

# Replace
data$reproductive_condition[x] <- new_values
```

Sort out remaining FIP info.

```{r}
x <- data$reproductive_condition == "no Fat In Pouch"

data[x,"fat_in_pouch"] <- "no FIP"
data[x,"reproductive_condition"] <- ""
```

```{r}
# "lpd" - going to assume this means "ldp"
index <- data$reproductive_condition == "lpd"
data[index, "pouch_condition"] <- "ldp"
data[index, "reproductive_condition"] <- ""

# "recent" values refere to post lactating recent. Remove these.
x <- data$reproductive_condition != ""
new_values <- data[x,"reproductive_condition"] %>%
  str_replace("recent, ", "") %>%
  str_replace("No", "no")
new_values %>% unique
```

Manually sort out the remaining values

```{r}
# What values?
data[,"reproductive_condition"] %>% unique

# "teets prom but no milky"
index <- data$reproductive_condition == "teets prom but no milky"
stopifnot(data[index, "nipple_condition"] == "")
data[index, "nipple_condition"] <- "swn, no milk"
data[index, "reproductive_condition"] <- ""

# "recent" - this can be removed since it refers to recent PL
data[data$reproductive_condition == "recent", "reproductive_condition"] <- ""

# "abort"
index <- data$reproductive_condition == "abort"
data[index, "other"] <- 
  append_note("aborted", data[index, "other"])
data[index, "reproductive_condition"] <- ""

# "not bred"
index <- data$reproductive_condition == "not bred"
data[index, "misc_reproductive_notes"] <- 
  append_note("not_bred", data[index, "misc_reproductive_notes"])
data[index, "reproductive_condition"] <- ""

# "0py"
index <- data$reproductive_condition == "0py"
stopifnot(data[index, "number_of_pouch_young"] == "0")
data[index, "reproductive_condition"] <- ""

# "m??"
index <- data$reproductive_condition == "m??"
data[index, "reproductive_condition"] <- ""

# "check; 16"
index <- data$reproductive_condition == "check; 16"
data[index, "pouch_young_size_in_mm"] <- "16"
data[index, "reproductive_condition"] <- ""

# "check; 25"
index <- data$reproductive_condition == "check; 25"
data[index, "pouch_young_size_in_mm"] <- "25"
data[index, "reproductive_condition"] <- ""

# "4 LN little or no milk"
index <- data$reproductive_condition == "4 LN little or no milk"
data[index, "nipple_condition"] <- "4ln, no milk"
data[index, "reproductive_condition"] <- ""

# "pos"
index <- data$reproductive_condition == "pos"
stopifnot(data[index,"misc_reproductive_notes"] == "pos bred in prev years")
data[index, "reproductive_condition"] <- ""

# "2 tags (8774, 8851)" - info already recorded in replaced_tags
index <- data$reproductive_condition == "2 tags (8774, 8851)"
stopifnot(replaced_tags[["8774"]] == c("8774", "8851"))
stopifnot(data[index,"other"] == "")
data[index,"other"] <- "Replaced tag 8851"
data[index, "reproductive_condition"] <- ""

# "may give birth soon"
index <- data$reproductive_condition == "may give birth soon"
data[index,"misc_reproductive_notes"] <-
  append_note("may give birth soon", data[index,"misc_reproductive_notes"])
data[index, "reproductive_condition"] <- ""
```

Check all values are empty.

```{r}
stopifnot(data$reproductive_condition == "")
```

-----

# Tidy columns

```{r}
colnames(data)
```

```{r}
# Remove unneeded columns and rename columns
tidy_data <- data %>% select(-day, -line, -point, -age, -py, -reproductive_condition, 
                -new, -first_weight) %>%
  rename(rep_con_string=rep_con, other_notes=other)

# Re-order columns
tidy_data <- tidy_data %>% select(
  date, location, id, sex, life_stage, weight, head, body, tail,
  pregnancy_status, pouch_condition, fat_in_pouch, number_of_pouch_young, 
  pouch_young_size_in_mm, lactation_status, nipple_condition, 
  testis_size_in_mm, misc_reproductive_notes,
  other_notes,
  country, state_or_region, mountain, year, 
  wild_captive, source_population, 
  taxon_id, phylum:genus, species, subspecies, common_name,
  identified_by, collector,
  everything())
```

-----

# Column classes

Change column classes if necessary.

```{r}
tidy_data %>% str
microsat_data %>% str
replaced_tags %>% str
```

```{r}
# Check that the only characters used are numeric
stopifnot(! str_detect(tidy_data$pouch_young_size_in_mm, "[^0-9.]"))
stopifnot(is.na(tidy_data$testis_size_in_mm) |
            ! str_detect(tidy_data$testis_size_in_mm, "[^0-9.]"))

# Change from characters to numeric
tidy_data <- tidy_data %>% 
  mutate(pouch_young_size_in_mm=as.numeric(pouch_young_size_in_mm),
         testis_size_in_mm=as.numeric(testis_size_in_mm))
```

```{r}
# Microsat data should be numeric
microsat_data <- microsat_data %>% 
  mutate_at(vars(BP1_FAM_allele1:BC36_PET_allele2), as.numeric)
```

```{r}
# View column names for tidy data
colnames(tidy_data)
```

```{r include=FALSE, eval=FALSE}
tidy_data %>% select(rep_con_string, pregnancy_status:misc_reproductive_notes) %>%
  View
```

-----

# Write output

Save data frames.

```{r}
saveRDS(tidy_data, file="cache/tidy_data.rds")
saveRDS(microsat_data, file="cache/microsat_data.rds")
saveRDS(replaced_tags, file="cache/replaced_tags.rds")
```

-----

# Session Info

```{r}
devtools::session_info()
```

